---
title: "Homework 3 Solutions"
author: "Matthew Ma"
date: "2022-10-16"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.


```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

ext is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digit=2)
```

### Problem 2

### Load tidy and wrangle the data

```{r}
acc_df = 
  read_csv("data/accel_data.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minutes",
    values_to = "activity_count"
  ) %>% 
  mutate( minutes = str_sub(minutes,10),
          minutes = as.numeric(minutes),
          day = forcats::fct_relevel(day, "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"),
          week_days = ifelse(day == "Saturday" | day == "Sunday", "weekend", "weekday")) %>% 
  select(week, day_id, day, week_days, everything()) %>% 
  arrange(week, day)
```

* The dataset `acc_df` has `r nrow(acc_df)` observations and `r ncol(acc_df)` variables with the names `r names(acc_df)`

```{r}
acc_df %>% 
  group_by(week, day) %>% 
  summarize(sum_activity = sum(activity_count, na.rm = TRUE)) %>%
  pivot_wider(
    names_from = "day",
    values_from = "sum_activity"
  ) %>% 
  knitr::kable(digits = 2)
```

### Making a plot to distinguish a trend

```{r}
acc_df %>% 
  group_by(week, day) %>% 
  summarize(sum_activity = sum(activity_count, na.rm = TRUE)) %>% 
  ggplot(aes(x = day, y = sum_activity, group = week, color = week)) +
  geom_point() + 
  geom_line()
```

* There is no apparent trend to be spotted on the graph. However, there are extremely low values spotted on the weekend compared to weekday. 

### Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week.

```{r}
acc_df %>% 
  mutate(hour = minutes / 60) %>% 
  ggplot(aes(x = hour, y = activity_count, color = day)) + 
  geom_line(alpha = 1)
```

* From this graph, we can conclude that there is a low activity count at night roughly from 11pm to 6am in the morning. The patient is most likely sleeping during the time period which explains the low activity count. And the activity starts to rise up after 6 am. Following a stable curve to rise up to the highest when it gets close to noon. Then there is a fluctuation between 12pm to 7pm. The activity count is the highest before the patient goes to sleep.

### Problem 3

```{r}
data("ny_noaa")

ny_df = 
  ny_noaa %>% 
  as_tibble(ny_noaa)

summary(ny_df)
```

```{r}

ny_df = 
  ny_noaa %>% 
  janitor::clean_names() %>%
  separate(col = date, into = c('year', 'month', 'day'), sep = "-", convert = TRUE) %>%
  mutate(
    month = month.name[month],
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    tmax = tmax/10,
    tmin = tmin/10,
    prcp = prcp/10)

ny_df %>% 
  count(snow, name = "n_obs") %>% 
  arrange(desc(n_obs))
```

* The most commonly observed value for snow is 0 because snow days are rare in New York which only happen during winter. Most of the other time there is no snow.

### Make a two panel plot showing the average max temperature of Janurary and July

```{r}

ny_df %>%
  group_by(id, year, month) %>% 
  filter(month == c("January", "July")) %>% 
  summarize(avg_tmax = mean(tmax)) %>% 
  ggplot(aes(x = year, y = avg_tmax, color = month)) +
  geom_point() +
  facet_grid(. ~ month) +
  labs( x = "Year",
        y = "Average Max Temperature",
        title = "Average Max Temperature in Janurary and July Throughout Years")
```

* There is a clear difference in the average max temperature between January and July from 1980 to 2010. The average max temperature of January is clearly lower than that of July. There are outliers in 1982 January and 1988 July. For both months the average max temperature fluctuate. 

### Make a plot showing tmax and tmin, make a plot showing the snow values greater than 0 and less than 100 in a year (separately).

```{r}
tmax_vs_tmin = 
  ny_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() +
  labs( x = "Minimum Temperature in Celsiusc",
        y = "Maximum Temperature in Celsius",
        title = " Tmax vs Tmin")

snowfall = 
  ny_df %>% 
  filter(snow > 0 & snow < 100) %>% 
  mutate(
    year = as.factor(year)
  ) %>% 
  ggplot(aes(x = snow, y = year)) +
  geom_density_ridges(scale = .5) + 
  theme(legend.position = "none") +
  labs( x = "Snowfall in mm",
        y = "Years",
        title = "Snowfall Throughout the Years")

tmax_vs_tmin + snowfall
```

